---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit the README.Rmd file -->


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lab report #1

Follow the instructions posted at https://ds202-at-isu.github.io/labs.html for the lab assignment. The work is meant to be finished during the lab time, but you have time until Monday evening to polish things.

Include your answers in this document (Rmd file). Make sure that it knits properly (into the md file). Upload both the Rmd and the md file to your repository.  

All submissions to the github repo will be automatically uploaded for grading once the due date is passed.
Submit a link to your repository on Canvas (only one submission per team) to signal to the instructors that you are done with your submission.

1. Inspect the first few lines of the dataset:

- what variables are there?
  - ParcelID, Address, Style, Occupancy, Sale Date, Sale Price, Multi Sale, Year Built, Acres, Total Living Area (ft^2), Bedrooms, Finished Basement (ft^2), Lot Area (ft^2), AC, Fire place, Neighborhood
  
- of what type are the variables?
  - chr, chr, factor with 12 levels, factor with 5 levels, date, num, chr, num, num, num, num, num, num, chr, chr, factor with 42 levels

- what does each variable mean? 
  - The majority of these variables are fairly self explantory such as parcel ID, address, sale date, sale price, Year built, acres, total living area, bedrooms, finished basement, lot area, AC, and fireplace.
  - The variables that are unique include Style which details the type of housing with 12 different types listed. Occupancy which details the type of housing that this is. Multi sale which determines whether the house was sold as a package deal. And finally Neighborhood which indiciates which neighborhood this house is a part of.

- what do we expect their data range to be?
  - The data ranges for the variables are as follows:

  - Style: 1-12 which correlate with style factor

  - Occupancy: 1-5 which correlate with occupancy factor

  - Sale Date: Date between Jan 1 2017 and Decemeber 31 2022

  - Sale Price: Number between 0 and 20500000

  - Year Built: num between 0 and 2022

  - Acres: Number between 0 and 13

  - Total Living Area: Number between 0 and 6007

  - Bedrooms: Number between 0 and 10

  - Finished Basement Area: Number between 10 and 6496

  - Lot Area: Number between 0 and 523228

  - Neighborhood: 1-42 which correlates with Neighborhood factor

2. is there a variable of special interest or focus?
    - The obvious variable of special interest is Sale Price; we chose this because all other variables have the potential to affect the Sale Price variable and it would likely have the highest range.

3. start the exploration with the main variable:

- what is the range of the variable? 
  - The range is between 0-20,500,000
- draw a histogram for a numeric vaiable or a bar chart, if the variable is categorical. 
```{r, message=FALSE}
library(ggplot2)
library(classdata)
ggplot(ames, aes(x=ames$"Sale Price")) + geom_histogram(binwidth = 6000) + xlim(0, 1100000) + ylim(0, 150)
```

- what is the general pattern? is there anything odd?
  - The distribution of Sale Price follows a right-skewed distribution around a median of 170900. 
  - There are several large spikes, including at 0, which are points in the data set where units of a building involved in a multisale were all marked as being sold for the total price of the apartment building. There were also many records with a recorded sale price of $0.

4. pick a variable that might be related to the main variable.

- what is the range of that variable? plot. describe the pattern.

- what is the relationship to the main variable? plot a scatterplot, boxplot or facetted barcharts (dependening on the types of variables involved). Describe overall pattern, does this variable describe any oddities discovered in 3? Identify/follow-up on any oddities.

Carter Parks: **Acres**- This variable has range 0 - 12.012
```{r}
ggplot(ames, aes(x = Acres)) +
  geom_histogram(binwidth = 0.1) +
  xlim(0, max(ames$Acres)) +
  ylim(0, 150) +
  labs(title = "Histogram of Acres",
       x = "Acres",
       y = "Frequency")
```

The acres variable follows an interesting pattern; most entries have values between 0 and 1, with the most entries in the first bin (0 - 0.1). Entries with higher values are few and far between; but there is one entry at 12 acres. 

```{r}
ggplot(ames, aes(x = Acres, y = `Sale Price`)) +
  geom_point(color = "blue") +
  labs(title = "Scatterplot of Acres vs Sale Price",
       x = "Acres",
       y = "Sale Price")
```

The acres variable has an odd relationship with Sale Price. It seems the vast majority of entries fall into the category with sub-2.5 acres and sales price less than 2,000,000. This makes it easy to identify the outliers, but more difficult to recognize a relationship between acres and sales price.

Ryan Rockey: **YearBuilt**


- This variable has the range 1880 - 2022. There is one record with a incorrect YearBuilt value of 0 which I am ignoring for my purposes.

```{r, message=FALSE}
ames_revised <- ames[ames$`YearBuilt` != 0, ] # removing the value where YearBuilt is 0

ggplot(ames_revised, aes(x="", y=YearBuilt)) + geom_boxplot() + ggtitle("Distribution of YearBuilt")
```

- The distribution of YearBuilt is heavily skewed left, which makes sense considering population growth will lead to more construction being required as time progresses.


``` {r, message=FALSE}
ggplot(ames_revised, aes(x=YearBuilt, y=`Sale Price`)) + geom_point() + scale_y_log10() + ggtitle("YearBuilt vs. Sale Price")
```

- The pattern of the scatter plot follows a slightly positive, linear correlation. Many outliers are present across all years.  
- One oddity the clear line of points across the chart where Sale Price = 0.  
- Also, because the dataset compiles properties sold from 2017 to 2022, there is a clear increase in properties after 2017 due to all the new constructions that are also accounted for.

Carolyn Jones: **AC**

- Because AC is a categorical variable, there is no range unless you map the responses to a number (0 for "No" and 1 for "Yes); doing this would give a range from 0-1.

```{r echo=TRUE}
ggplot(ames, aes(x = AC)) + geom_bar(fill = 'pink')
```

- There isn't too much to explain here since the graph is very simple. Our variable AC has 2 different categories/values, "Yes" and "No". From the graph we can tell that a majority of properties in Ames have AC, with the number of "Yes" answers being around 6 times higher than the number of "No" answers.

```{r echo=TRUE}
ames2 <- ames[ames$`Sale Price` != 0,]
ggplot(ames2, aes(x = AC, y = `Sale Price`, color = AC)) + geom_boxplot() + scale_y_continuous(limits = c(0, 500000))
```

- From the box plot graph above, we can see a clear correlation between the sale price of a property and the presence of AC. The IQR's of both plots are relatively the same size, however the values that make them up differ drastically. The first quartile for properties with AC is around the same value as the third quartile for properties without; this shows that values with AC typically sell for much higher than values without. One oddity to note about the graph is the number of outliers in the "No" boxplot. 


Ryan Ledbetter **Bedrooms**

- The rang fro bedrooms is from 0 to 10. There are a few that use N/A but these will be ignored as I am treating them as incomplete data.
```{r}
ggplot(ames, aes(x = Bedrooms)) + geom_histogram(binwidth = 1) + xlim(0, 10) + ylim(0, 2500) + labs(title = "Histogram of Bedrooms", x = "Number of bedrooms", y = "Frequency")
```

- The distribution of the number of bedrooms for houses in Ames is a normal distribution that is skew right.

```{r}
ames3 <- ames[ames$Bedrooms != 'N/A', ]
ames3 <- ames3[ames3$`Sale Price` != 0,]
ames3 <- ames3[ames3$`Sale Price` < 14200000,]
ames3$Bedrooms <- as.factor(ames3$Bedrooms)
ggplot(ames3,aes(x = Bedrooms, y = `Sale Price`)) + geom_boxplot() + labs(title = "Box plot of Bedrooms vs Sale Price", x = "Number of bedrooms", y = "Sale Price")
```
- For this boxplot, I removed the two highest prices for sales as they were a part of the multi-sale and did not fully represent the data. What can be seen from the box plot here, is that as the number for bedrooms increases, on average the price also increases. The seven bedroom houses and the ten bedroom houses act as outliers in this data set as they do not follow the  trend of the rest of the graph, but this could be explained by lacking amounts of data seen from the histogram.

- For the sake of completeness, I have included a boxplot showing the data including the highest priced houses that were sold in the large multisale.
```{r}
ames3 <- ames[ames$Bedrooms != 'N/A', ]
ames3 <- ames3[ames3$`Sale Price` != 0,]
ames3$Bedrooms <- as.factor(ames3$Bedrooms)
ggplot(ames3,aes(x = Bedrooms, y = `Sale Price`)) + geom_boxplot() + labs(title = "Box plot of Bedrooms vs Sale Price", x = "Number of bedrooms", y = "Sale Price")
```

